\section{Objectives}
\label{s:objectives}
\hspace{10pt} This project aims to investigate the energy efficiency of phone sensing. It sets to review the current state of the art and identify possible enhancements. I intent to establish energy efficiency levels of different sensors in the experiments. The results of those experiments are planned to be leveraged for the design of energy-efficient sensing library. This library should be implemented and evaluated in real-life scenarios. More specifically, the project has the following set of objectives: 

\begin{enumerate}
 \item Primary objectives
  \begin{enumerate}
  	\item \label{o:survey:review}To review the current state of the art in sensing, energy measurement methods and energy-efficient sensing.
  	\item \label{o:survey:enhancements}To identify possible enhancements in energy measurement methods and energy-efficient sensing.
  	\item \label{o:measurements:primary}To determine energy efficiency levels of different sensors.
  	\item \label{o:library:design}To design the energy-efficient library, which leverages established energy efficiency levels.
  	\item \label{o:library:impleval}To implement the energy-efficient library and evaluate it in real-life scenarios.
  \end{enumerate}
  \item Secondary objectives
  \begin{enumerate}
    \item \label{o:measurements:empirically}To introduce a novel method for energy measurement and validate it empirically.
    \item \label{o:measurements:online}To investigate the possibility of online energy measurement.
    \item \label{o:measurements:differentphones}To establish and compare sensors' energy efficiency levels among different mobile phones.
    \item \label{o:library:battery}To adapt energy saving algorithm depending on a current battery life.
  	\item \label{o:library:devices}To propose device-independent energy-efficient strategy for phone sensing.
  \end{enumerate}
  \item Tertiary objectives
  \begin{enumerate}
   \item \label{o:library:original}To propose an original technique for energy-efficient sensing.
  \end{enumerate}
\end{enumerate}