\section{Software engineering processes}
\label{s:processes}
\hspace{10pt} Beside Dr Tristan Henderson being my supervisor, he also acted as \textbf{a fictional customer} of my software product. The continuous contact with the client allowed us to apply \textbf{Agile} approach to the project. At the beginning of our cooperation,  we established what we want to achieve (objectives) and how to measure whether we have achieved them (requirements). During the project, we used those requirements to indicate whether we are on right track to finish the product on time.

We applied \textbf{Test-Driven Development}: every week I tested and implemented a next block of the final product. Then, I document my work's progress\ (all progress files are available in \textit{progress/} directory) and meet with Tristan. He provided me with continuous feedback on my work and assess whether I met last week's aims. At the end of each of the meetings, we set aims for the next week. This approach was suitable for \textbf{a research project}, as it was often difficult to estimate time, which I need to complete different parts of the project. Some of our objectives were never done before by research community and we simply did not even know whether they were possible. The subject of energy-efficient sensing is very complex and if we had chosen different software engineering process, there would be a risk of not delivering the product on time. By making short development cycles and having continuous feedback, I could achieve the objectives we set at the beginning. For example, Locy went through a couple of phases. First, there were two applications: an application continuously sampling GPS\ (baseline application) and another application continuously sampling accelerometer. The energy efficieny of both applications were compared. The latter application had lower energy demands, and thus, it was evolved into its new form. The new form of the application inferred the movement detection and switched on GPS when a user was not walking. After it was implemented, its energy efficiency was evaluated. The result was not so promising, and thus, the duty-cycling was involved into the application. This was followed by another series of energy efficiency experiments. Finally, the application was changed into separate library called Locy. Such development cycle shows how early testing (evaluating energy efficiency ) has affected the design decisions of the project.
			
For the development process, we also used a variety of support software. The code was hosted and shared between myself and Tristan by using popular source code management tool, Git. From the very beginning, the whole project\ (except for Tristan's ESM) was open-sourced and publicly available on GitHub: \url{https://github.com/mk74/locy}. We believed that this would provide higher quality of code and attract other researchers, who work on the same problem. All applications were created in Android Developer Tools \cite{google:adt}, which is a special plugin for Eclipse. Google provides the new IDE for Android, Android Studio \cite{google:androidstudio}. Although It has an interesting features specifically for Android development e.g., code refactoring to run faster, Android Studio is still in beta version. The testing of our library project involved energy measurement, which we couldn't automate and thus why none of testing framework was used. Lastly, we used CiteULike for sharing research papers and management of references for this document.
	